---
title: "Acitivity Monitoring"
author: "Tuong Nguyen"
date: "October 12, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('dplyr')
library('data.table')
library('ggplot2')
```

## Loading and preprocessing the data
```{r echo=FALSE, results='markup'}
activity <- read.csv('activity.csv')
head(activity)

```

## What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day
2. Make a histogram of the total number of steps taken each day
3. Calculate and report the mean and median of the total number of steps taken per day
```{r, echo=FALSE }
daily <- aggregate(x = activity$steps, by = list(activity$date), sum, na.rm = TRUE)
colnames(daily) <- c("date","steps")
hist(daily$steps,breaks = 50, 
     main = "Histogram of the total number of steps taken daily",
     xlab = 'number of steps')

dailymean <- round(mean(daily$steps),0)
dailymedian <- median(daily$steps)
abline(v = dailymean, col = 'red', lty = 1, lwd = 3)
abline(v = dailymedian, col = 'blue', lty = 1, lwd = 3)

```


#### Mean of the total of steps taken per day is `r dailymean` (red vertical line)


#### Median of the total of steps taken per day is `r dailymedian` (blue vertical line)


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
